%
\documentclass[10pt]{exam2}
%\printanswers

\usepackage{amsmath}

\begin{document}

%\maketitle
\begin{center}
{\LARGE Quiz 1 - Foundations}\\\vspace{2mm}
\vspace{3mm}
{\large Advanced Methods}\\\vspace{2mm}
%Carlisle Rainey\symbolfootnote[1]{Carlisle Rainey is an Assistant Professor of Political Science, University at Buffalo, SUNY, 520 Park Hall, Buffalo, NY 14260 (\href{mailto:rcrainey@buffalo.edu}{rcrainey@buffalo.edu}).}
\end{center}

\begin{questions}
\question Define to matrices $A$ and $B$, such that 
\begin{equation}\nonumber
A = \begin{bmatrix}
       1 & 2           \\
       3 & 4
     \end{bmatrix}
\end{equation}
and 
\begin{equation}
B = \begin{bmatrix}\nonumber
       1 & 2 & 3           \\
       4 & 5 & 6
     \end{bmatrix}~\text{.}
\end{equation}

\begin{parts}
\part What are the dimensions of $B$? 
\part What is $AB$?
\part What is $BA$?
\end{parts}

\question Find the first derivatives.
\begin{parts}
\part $f(x) = (x^2 - 7)^{10}$
\part $f(x) = e^{4x}$
\part $f(x) = (6 - x)^9\text{ln}(x^2)$
\end{parts}

\question Find the indefinite integrals.
\begin{parts}
\part $f(x) = 4$
\part $f(x) = 3x^2 - 2x$
\part $g(z) = e^{4z}$
\end{parts}

\question For $g(z)$ above, find $\int_0^1g(z) dz$.

\question In the last class, Kevin suggested that two events $A$ and $B$ are independent \textit{if and only if} $P(A) = P(A|B)$. Instead, I suggested the two events are independent if $P(AB) = P(A)P(B)$. I told him my definition implies his, but I wasn't sure if his implied mine. Kevin was right. Prove that $P(A) = P(A|B) \iff P(AB) = P(A)P(B)$. Note that an \textit{if and only if} proof requires both directions.
\begin{parts}
\part Prove that $P(AB) = P(A)P(B) \Rightarrow  P(A) = P(A|B)$. (I was sure about this one.)
\part Prove that $P(A) = P(A|B) \Rightarrow P(AB) = P(A)P(B)$. (I wasn't sure about this one.)
\end{parts}

\question Suppose that African-Americans account for 55\% of violent crimes in the U.S. and make up 13\% of the population. Suppose that about 0.5\% of adults in the U.S. are violent criminals. Richard Cohen has argued that it is not racist to suspect people like Trayvon Martin of being a violent criminal, because the disproportional share of violent crimes are committed by African-Americans. If you encountered a random African-American, what is the probability that person is a violent criminal?

\question Suppose the pdf $f(x) = 3x^2$ for $0 \leq x \leq 1$.
\begin{parts}
\part Show that this is indeed a pdf.
\part What is the probability of $X > \dfrac{1}{2}$.
\end{parts} 

\end{questions}
\end{document}





